The challenge and promise of software citation for credit, identification, discovery, and reuse

### 2016-04-10 response-to-reviewers.tex

\documentclass{article}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx, caption}
\usepackage{latexsym,url,amscd, amssymb, amsmath}
\usepackage{epstopdf}

\usepackage{multicol}
\usepackage{booktabs}
\usepackage{fancyvrb}
\usepackage{url,breakurl}

\usepackage[english]{babel}
\usepackage{textcomp}

%to highlight text
\usepackage[usenames, dvipsnames]{color}
\providecommand{\hly}[1]{\colorbox{yellow}{#1}}
%\providecommand{\hlt}[1]{\colorbox{Turquoise}{#1}}
%\providecommand{\hlg}[1]{\colorbox{green}{#1}}
\providecommand{\hlr}[1]{\colorbox{SkyBlue}{#1}}

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{}{;}
\usepackage[linkcolor=blue, citecolor=blue, colorlinks=true, breaklinks=true]{hyperref}

\usepackage{setspace}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%\date{}                                           % Activate to display a given date or no date

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.4cm}

%%% BEGIN DOCUMENT
\begin{document}

We thank both reviewers for their detailed comments, and have addressed their requests and suggestions here.
We performed a significant revision of our manuscript.
%Specific changes in the marked revision in response to Reviewers \#1 and \#3 are respectively highlighted in \hly{yellow} and \hlr{blue}.
%Reviewer \#2's comments were favorable, and no action was required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer \#1:}

\textbf{Reviewer comment:}
\emph{This is an interesting topic and authors highlight important concerns about software citation and summarize the scope of the problem providing, however, few possible answers or a clear path that might be followed.}

\textbf{Our response:}
We added a table that summarizes the key challenges and possible solutions to \textbf{page 3}.

\textbf{Reviewer comment:}
\emph{Also, the discussion can go deeper other than ways to cite software. In cases where software is not the focus of the research should it even be cited? For instance, in a data analysis scenario, does it makes sense to cite a particular implementation of an algorithm? Should Pearson be cited in the references section when a correlation analysis is performed?}

\textbf{Our response:}
We agree that this is an important point, and have added discussion of when software should be cited to \textbf{page 4}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Reviewer \#2:}

\textbf{Reviewer comment:}
\emph{One method to increase\ldots who suggests this? (yourself, or should there be a reference here).}

\textbf{Our response:}
We have added greater discussion and references for this statement to \textbf{page 1}.

\textbf{Reviewer comment:}
\emph{Software citation is indeed as important as article or data citation. The connection with DQ\slash IQ seems a bit tenuous, however. I am sure the connection can be made, for instance the ability to do all the things promised in the title relies upon good quality and complete metadata.}

\emph{However I was rather hoping the authors would elaborate on this connection?}

\textbf{Our response:}
We have added discussion and references on the impact of software citation to DQ\slash IQ through reproducibility and provenance of research results on \textbf{pages 1--2}.

\textbf{Reviewer comment:}
\emph{I am aware of at least one recent workshop on software citation (organised at the University of Manchester in the UK), so the topic is clearly on the table. On this note, I would suggest the authors may want to frame the related work discussion in the context of the challenges, as it seems to be work (for the most part, it's not research) aimed at meeting some of those challenges.}

\textbf{Our response:}
We added discussion and references to multiple recent workshops on software citation to \textbf{page 2}.
In addition, the entire discussion of related work on \textbf{pages 2--3} parallels the key research challenges summarized in Table 1 on \textbf{page 3}.

\textbf{Reviewer comment:}
\emph{There are 5 bibliography entries in the references (i.e., proper publications), and 4 are self-citations. I find this rather unbalanced.}

\textbf{Our response:}
Through the course of extending discussion and adding content in response to other comments, the bibliography has grown significantly.
Out of the 36 total entries, 18 are to ``proper'' publications (i.e., journal articles, conference proceedings, or published technical reports), and one or more of us contributed to only four of those.

\end{document}

### 2016-06-26 SC-challenge.tex
removed outdated quote on what software to cite

In arguing the need for software citation, we simultaneously introduce the question of \textbf{when} software should be cited.
Although there is no clear answer to this question, we argue that ensuring reproducibility of research results requires citation of software if used directly and important to research results.
In other words, if using different software could produce different data or results, then the software used should be cited.
We are currently leading the FORCE11 Software Citation Working Group to develop principles for software citation~\citeyear{SCWG}.
The draft principles document~\cite{SoftwareCitation2016}, which is adapted from an earlier set of principles for data citation~\cite{DataCitation2014}, discusses how to decide what software to cite:
\begin{quote}
What software should be cited is the decision of the author(s) of the research work in the context of community norms and practices, and in most research communities, these are currently in flux.
In general, we believe that software should be cited on the same basis as any other research product such as a paper or book; that is, authors should cite the appropriate set of software products just as they cite the appropriate set of papers.
Again, the specific decision of what is appropriate must be made by the author(s) of the product.
However, an illustrative example is the use of Microsoft Excel in research.
We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.  Similarly, general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), research presentations (e.g., Powerpoint) or communications (e.g., Skype) should not be cited.
This recommendation matches that of the Purdue Online Writing Lab: ``Do not cite standard office software (e.g. Word, Excel) or programming languages.  Provide references only for specialized software''~\cite{powl-citing-software}.
In other words, if using different software could produce different data or results, then the software used should be cited.
%
%Note that some software which is or could be captured as part of data provenance may not be cited.
%Citation is a record of software that is important to the research outcome, where provenance is a record of all steps (including software) used to generated particular data within the research process.
%This implies that for a data research product, provenance data will include all cited software, but not necessarily vice versa.
%Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.
%And the data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
%These statements may also be true for software products.
%In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
%Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
\end{quote}

>>>>

In arguing the need for software citation, we simultaneously introduce the question of \textbf{when} software should be cited.
Although there is no clear answer to this question, ensuring reproducibility of research results requires citation of software if used directly and important to research results.
In other words, if using different software could produce different data or results, then the software used should be cited.
We led the FORCE11 Software Citation Working Group to develop principles for software citation~\cite{SoftwareCitation2016}, which are adapted from an earlier set of principles for data citation~\cite{DataCitation2014}; that document discusses further how to decide what software to cite.
